---
title: "Coursera Practical Machine Learning Course Final Project"
author: "Anna_Kozina"
date: "13 01 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load packages}
  
  if(!require(dplyr)){
    install.packages("dplyr", lib="/dplyr/")
    library(dplyr)
  }

  if(!require(lattice)){
    install.packages("lattice", lib="/lattice/")
    library(lattice)
  }
  
  
  if(!require(caret)){
    install.packages("caret", lib="/caret/")
    library(caret)
  }
  
  if(!require(rpart)){
    install.packages("rpart", lib="/rpart/")
    library(lubridate)
  }
  
  if(!require(rfm)){
    install.packages("rpart.plot", lib="/rpart.plot/")
    library(lubridate)
  }
  
  if(!require(RColorBrewer)){
    install.packages("RColorBrewer", lib="/RColorBrewer/")
    library(RColorBrewer)
  }
  
  if(!require(rattle)){
    install.packages("rattle", lib="/rattle/")
    library(rattle)
  }
  
  if(!require(e1071)){
    install.packages("e1071", lib="/e1071/")
    library(e1071)
  }
  
  if(!require(randomForest)){
    install.packages("randomForest", lib="/randomForest/")
    library(randomForest)
  }
  
  if(!require(RCurl)){
    install.packages("RCurl", lib="/RCurl/")
    library(RCurl)
  }
  
```

```{r download data}

  train <- getURL("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",
                ssl.verifyhost=FALSE, ssl.verifypeer=FALSE)
  
  train <- read.csv(textConnection(train), header=T, na.strings=c("NA","#DIV/0!",""))


  
  test <- getURL("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",
               ssl.verifyhost=FALSE, ssl.verifypeer=FALSE)
  
  test <- read.csv(textConnection(test), header=T, na.strings=c("NA","#DIV/0!",""))

  
  train  <- train[, colSums(is.na(train)) == 0]

  str(train)
```  
  
  
Check for near zero variance predictors and drop them if necessary
  
```{r near zero variance}

  nzv <- nearZeroVar(train)
  train <- train[, -nzv]
  dim(train)
  
```  
  
Split data

```{r split}

names(train)
str(train)
summary(train$classe)

library(caret)
set.seed(123)
DataPartition <- createDataPartition(y=train$classe, p=0.7, list=FALSE)
train_DataPartition <- train[DataPartition, ]
test_DataPartition <- train[-DataPartition, ]

str(train_DataPartition)
table(train_DataPartition$classe) 
```
  
Build random Forest
 
```{r random forest}

  library(randomForest)
  set.seed(222)
  rf <- randomForest(classe~., data=train)
  print(rf)

```
    
Predict (train data)

```{r predict}

  library(caret)
  predict <- predict(rf, train_DataPartition)
  table(predict)
  confusionMatrix(predict, train_DataPartition$classe)

```
  
  
  
The out-of-bag error 
    
There is no need for cross-validation/eparate test set to get an unbiased estimate of the test set error.   It is estimated internally, during the run, as follows:
Each tree is constructed using a different bootstrap sample from the original data. 
About 1/3 of the cases are left out of the bootstrap sample and not used in the construction of the
tree.
  
Put each case left out in the construction of the kth tree down the kth tree to get a classification. 
In this way, a test set classification is obtained for each case in about one-third of the trees. 
At the end of the run, take j to be the class that got most of the votes every time case n was oob. 
The proportion of times that j is not equal to the true class 
of n averaged over all cases is the oob error estimate. This has proven to be unbiased in many tests
  
Predict (test data)

```{r predict test}
  library(caret)
  
  predict_2 <- predict(rf, test_DataPartition)
  table(predict_2)

  confusionMatrix(predict_2, test_DataPartition$classe)
  
```

Summary
  
As shown in the accuracy result above, when tested on the subset data used for 
cross validation we get an accuracy of 99%. This can be used to estimate our out of sample error. 
On the other hand the fabricated model is fitted to the training portion of the data, we get an accuracy 
of 100%. This can be used to estimate the in sample


